---
title: "Chi^2 prunning"
output: html_notebook
---
Data generation:
```{r}
samples = 2000
a1 = runif(samples, 0, 1) < 0.5;
a2 = runif(samples, 0, 1) < 0.5;
ans = xor(a1, a2);
ts = data.frame(a1, a2, ans);
colnames(ts) <- c("A1", "A2", "Ans");
ts
```
Some helper functions:
```{r}
# return the most likely answer when no further evidence is avaialable
getmode <- function(v) {
   length(v["Ans"] == TRUE) > length(v["Ans"])/2
}
```

```{r}
# gives entropy contained in a list of boolean values
booleanEntropy <- function(truth_table){
  tab = table(truth_table);
  if(! "TRUE" %in% names(tab) || ! "FALSE" %in% names(tab))
    return(0)
  pos <- table(truth_table)["FALSE"]
  names(pos) <- "bool entropy"
  p <- pos/length(truth_table) 
  return(- p * log2(p) - (1-p) * log2(1-p))
}
```

```{r}
# calculates information gain of splitting by given attribute
importance <- function(atr, examples){
  val1 = examples[examples[atr] == TRUE, ]
  val2 = examples[examples[atr] == FALSE, ]
  s_true <- booleanEntropy(val1[,"Ans"])
  s_false <- booleanEntropy(val2[, "Ans"])
  print(paste("Entropies:", s_true, s_false))
  R <- (s_false*length(val1[, "Ans"]) + 
          s_true*length(val2[, "Ans"]))/(length(examples[, atr]))
  names(R) <- paste("Gain on", atr)
  return(-R + booleanEntropy(examples[,"Ans"])) # strange ordering to preserve name
}
importance("Ans", ts)
```


Decision tree learning function:
```{r}
decisionTreeLearning <- function(examples, attributes, parent_examples){
  if (length(examples[,"Ans"]) == 0)
    return(getmode(parent_examples))
    # return most probable answer as there is no training data left
  else if( length(attributes) == 0)
    return(getmode(examples))
  else if (length(table(examples["Ans"]))) < 2):
    return(examples["Ans", 1])
  
  A = max(attributes, key(a)=importance(a, examples))
  # choose the most promissing attribute to condition on
  tree = new Tree(root=A)
  for( value in A.values()){
    exs = examples[e.A == value]
    subtree = decisionTreeLearning(exs, attributes.remove(A), examples)
    # note implementation should probably wrap the trivial case returns into trees for consistency
    tree.addSubtreeAsBranch(subtree, label=(A, value)
  }
  
  return(tree)
}
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
